---
title: "MAT 120 Computer Lab - Week 7"
subtitle: "Normal Distribution"
output:
  html_document: default
  pdf_document: default
---

# Learning goals

- standardize a value using a z-score,
- find normal probabilities,
- find percentiles and cutoff values,
- interpret normal-model results in context.

# Standard normal distribution

Area to the left of z = 1.80:

```{r}
pnorm(1.80)
```

Area to the right:

```{r}
1 - pnorm(1.80)
```

Middle probability:

```{r}
pnorm(1.5) - pnorm(-1.5)
```

Find the z-score at the 95th percentile:

```{r}
qnorm(0.95)
```

# Example: Race finishing times

Assume 10-km race times are normally distributed with:

- mean = 61 minutes
- standard deviation = 9 minutes

Probability a randomly selected runner takes more than 72 minutes:

```{r}
1 - pnorm(72, mean = 61, sd = 9)
```

Probability between 52 and 70 minutes:

```{r}
pnorm(70, mean = 61, sd = 9) -
  pnorm(52, mean = 61, sd = 9)
```

95th percentile:

```{r}
qnorm(0.95, mean = 61, sd = 9)
```

# Questions

1. Find the z-score corresponding to 72 minutes.
2. Interpret the probability above 72 minutes.
3. Interpret the 95th percentile in context.

# Your Turn

Suppose exam scores are normally distributed with mean 72 and standard deviation 8.

1. Find P(X > 80).
2. Find P(65 < X < 85).
3. Find the 90th percentile.
4. Find the cutoff for the top 15% of scores.
