---
title: "MAT 120 Computer Lab - Week 6"
subtitle: "Discrete Random Variables and Binomial Distribution"
output:
  html_document: default
  pdf_document: default
---

# Learning goals

- build a probability distribution,
- calculate expected value,
- identify a binomial experiment,
- calculate binomial probabilities using R.

# 1. A discrete probability distribution

A fair six-sided die is rolled once. If the result is odd, you win the number
of dollars shown. If the result is even, you lose that amount.

```{r}
x <- c(1,-2,3,-4,5,-6)
p <- rep(1/6, 6)

data.frame(Profit = x, Probability = p)
```

Expected value:

```{r}
sum(x * p)
```

# Questions

1. What is the expected profit/loss per game?
2. Interpret the expected value using the Law of Large Numbers.
3. About how much would you expect to win or lose after 100 games?

# 2. Binomial probability

Suppose 20% of customers choose express delivery. If 12 customers are selected,
let X be the number who choose express delivery.

For a binomial model:

- fixed number of trials: n = 12,
- two outcomes per trial,
- constant probability p = 0.20,
- trials are independent.

Probability exactly 3 choose express delivery:

```{r}
dbinom(3, size = 12, prob = 0.20)
```

Probability at most 3:

```{r}
pbinom(3, size = 12, prob = 0.20)
```

Probability at least 3:

```{r}
1 - pbinom(2, size = 12, prob = 0.20)
```

Mean and standard deviation:

```{r}
n <- 12
p_success <- 0.20

mu <- n * p_success
sigma <- sqrt(n * p_success * (1 - p_success))

mu
sigma
```

# Your Turn

A very large shipment contains 2% defective items. Ten items are randomly selected.

1. Explain why a binomial model is reasonable.
2. Find P(X = 0).
3. Find P(X <= 1).
4. Find the expected number of defective items.
5. Find the standard deviation.
