---
title: "MAT 120 Computer Lab - Week 1"
subtitle: "Sampling, Populations, Samples, and Variables"
output:
  html_document: default
  pdf_document: default
---

# Learning goals

By the end of this lab, you should be able to:

- distinguish a population from a sample,
- identify an individual and a variable,
- distinguish categorical and numerical variables,
- generate a simple random sample using R.

# 1. Getting started with R

Try a few calculations.

```{r}
2 + 5
10 / 4
sqrt(81)
```

Create a vector of student ID numbers from 1 to 30.

```{r}
student_id <- 1:30
student_id
```

# 2. Random sampling

Select 5 students at random.

```{r}
sample(student_id, 5)
```

Run the command several times. Your sample should usually change.

To make a result reproducible, set a random-number seed.

```{r}
set.seed(120)
sample(student_id, 5)
```

# 3. Population and sample example

Suppose a college has 300 students in an introductory statistics course.
A professor randomly selects 15 students and records whether each completed
the homework.

```{r}
homework <- c("Yes","Yes","No","Yes","Yes",
              "Yes","Yes","No","Yes","Yes",
              "Yes","Yes","Yes","No","Yes")
table(homework)
prop.table(table(homework))
```

## Questions

1. What is the population?
2. What is the sample?
3. What is the sample size?
4. What is the variable?
5. Is the variable categorical or numerical?
6. What proportion of sampled students completed the homework?
7. Is that proportion a statistic or a parameter?

# Your Turn

A university has 800 students enrolled in introductory statistics.
Use R to randomly choose 20 ID numbers from 1 through 800.

```{r}
# Write your code here
```

Explain why the selected 20 students form a **sample** rather than a population.
