---
title: "MAT 120 Computer Lab - Week 9"
subtitle: "Confidence Intervals"
output:
  html_document: default
  pdf_document: default
---

# Learning goals

- construct a one-sample t confidence interval,
- interpret a confidence interval correctly,
- calculate margin of error,
- understand how confidence level and sample size affect interval width.

# Example: Newborn weights

Suppose the weights (lb) of 10 randomly selected newborns are:

```{r}
weights <- c(6.1, 6.8, 7.0, 7.2, 7.4,
             7.5, 7.7, 7.8, 8.1, 7.0)
```

Sample statistics:

```{r}
mean(weights)
sd(weights)
length(weights)
```

90% confidence interval:

```{r}
t.test(weights, conf.level = 0.90)$conf.int
```

# Manual calculation

```{r}
xbar <- mean(weights)
s <- sd(weights)
n <- length(weights)
df <- n - 1

t_star <- qt(1 - 0.10/2, df)
margin_error <- t_star * s / sqrt(n)

lower <- xbar - margin_error
upper <- xbar + margin_error

t_star
margin_error
c(lower, upper)
```

# Questions

1. State the interval.
2. Interpret it in context.
3. What is the margin of error?
4. Would a 95% confidence interval be wider or narrower than a 90% interval using the same data?
5. What happens to margin of error when sample size increases?

# Sample-size planning when sigma is known

Suppose we want to estimate a population mean within E = 20 units,
using sigma = 63 and 95% confidence.

```{r}
z_star <- qnorm(0.975)
sigma <- 63
E <- 20

n_required <- (z_star * sigma / E)^2
n_required
ceiling(n_required)
```

# Your Turn

A sample of 40 grading times has:

- sample mean = 12.6 minutes
- sample SD = 2.6 minutes

Construct a 90% confidence interval manually using the t distribution.
Then explain whether the interval supports the claim that the population mean exceeds 11 minutes.
