---
title: "MAT 120 Computer Lab - Week 10"
subtitle: "One-Sample Hypothesis Testing"
output:
  html_document: default
  pdf_document: default
---

# Learning goals

- state null and alternative hypotheses,
- calculate a one-sample t statistic,
- obtain and interpret a p-value,
- make a statistical decision using alpha,
- write a conclusion in context.

# Example: Average wind speed

A researcher claims that the population mean wind speed is 8 mph.

A random sample of 32 days gives:

- x-bar = 8.2 mph
- s = 0.6 mph
- n = 32

We test:

H0: mu = 8

Ha: mu != 8

at alpha = 0.05.

# Test statistic

```{r}
xbar <- 8.2
mu0 <- 8
s <- 0.6
n <- 32

t_stat <- (xbar - mu0) / (s / sqrt(n))
t_stat
```

Degrees of freedom:

```{r}
df <- n - 1
df
```

Two-tailed p-value:

```{r}
p_value <- 2 * (1 - pt(abs(t_stat), df))
p_value
```

# Decision

```{r}
alpha <- 0.05

if (p_value < alpha) {
  print("Reject H0")
} else {
  print("Do not reject H0")
}
```

# Questions

1. State H0 and Ha.
2. Is this a left-tailed, right-tailed, or two-tailed test?
3. Report the test statistic.
4. Report the p-value.
5. Compare the p-value with alpha.
6. Write the final conclusion in context.

# Using raw data with t.test()

```{r}
visits <- c(3,2,1,3,7,2,9,4,6,6,
            8,0,5,6,4,2,1,3,4,1)

mean(visits)
sd(visits)

t.test(visits,
       mu = 5.8,
       alternative = "less")
```

# Your Turn

A bank reports that the historical mean savings account balance is $1300.
A new random sample of 45 accounts has:

- x-bar = $1350
- s = $80

At alpha = 0.10, test whether the current population mean balance is different from $1300.

Show:

1. H0 and Ha
2. test statistic
3. degrees of freedom
4. p-value
5. decision
6. conclusion in context
