---
title: "MAT 120 Computer Lab - Week 2"
subtitle: "Frequency Tables and Graphs"
output:
  html_document: default
  pdf_document: default
---

# Learning goals

- construct frequency and relative-frequency tables,
- create bar charts for categorical data,
- create histograms for quantitative data,
- recognize discrete and continuous variables.

# 1. Categorical data

A class survey asked students how they usually travel to campus.

```{r}
transport <- c(
  "Subway","Bus","Subway","Walk","Subway","Bus","Car","Subway",
  "Walk","Bus","Subway","Subway","Car","Bus","Walk","Subway",
  "Bus","Subway","Walk","Subway"
)

freq <- table(transport)
freq
```

Relative frequency:

```{r}
rel_freq <- prop.table(freq)
rel_freq
```

Bar chart:

```{r}
barplot(freq,
        main = "Transportation to Campus",
        xlab = "Transportation Method",
        ylab = "Frequency")
```

# 2. Quantitative data

The following values are the numbers of minutes 30 students spent commuting.

```{r}
commute <- c(
  18,22,25,27,29,30,31,33,35,35,
  36,38,39,40,41,42,43,45,46,48,
  50,52,54,55,58,60,62,67,72,80
)
```

Create a histogram.

```{r}
hist(commute,
     main = "Student Commute Times",
     xlab = "Commute Time (minutes)",
     ylab = "Frequency")
```

# Questions

1. What is the individual in the transportation example?
2. What is the variable?
3. Why is a bar chart appropriate?
4. What is the variable in the commute example?
5. Is commute time discrete or continuous?
6. Describe the approximate shape of the commute-time distribution.

# Your Turn

Use the data below.

```{r}
coffee <- c(12,18,21,15,25,24,17,19,22,28,
            11,16,20,23,27,14,18,22,26,30)
```

1. Make a histogram.
2. Find the minimum and maximum.
3. Estimate a reasonable class/bin width.
4. State whether the variable is discrete or continuous in context.
