---
title: "MAT 120 Computer Lab - Week 4"
subtitle: "Correlation and Linear Regression"
output:
  html_document: default
  pdf_document: default
---

# Learning goals

- create a scatterplot,
- describe direction, form, and strength,
- calculate a correlation coefficient,
- fit and interpret a least-squares regression line,
- recognize extrapolation.

# Example: Study time and exam score

```{r}
hours <- c(1,2,2,3,3,4,4,5,5,6,6,7)
score <- c(58,61,65,66,70,72,75,78,81,84,86,91)
```

Scatterplot:

```{r}
plot(hours, score,
     main = "Exam Score vs. Study Time",
     xlab = "Study Time (hours)",
     ylab = "Exam Score")
```

Correlation:

```{r}
cor(hours, score)
```

Regression model:

```{r}
fit <- lm(score ~ hours)
summary(fit)
```

Add the regression line:

```{r}
plot(hours, score,
     xlab = "Study Time (hours)",
     ylab = "Exam Score")
abline(fit)
```

Make a prediction for 5.5 hours.

```{r}
predict(fit, newdata = data.frame(hours = 5.5))
```

# Questions

1. Identify the explanatory variable.
2. Identify the response variable.
3. Describe the direction of association.
4. Is the association approximately linear?
5. Interpret the slope in context.
6. What does the sign of the correlation tell you?
7. Would predicting a score for 20 hours be reliable? Explain.

# Your Turn: Cars

R contains the built-in `mtcars` data set.

```{r}
head(mtcars)
```

Study the relationship between vehicle weight (`wt`) and fuel economy (`mpg`).

1. Make a scatterplot.
2. Calculate the correlation.
3. Fit `lm(mpg ~ wt)`.
4. Add the regression line.
5. Interpret the slope.
