---
title: "MAT 120 Computer Lab - Week 3"
subtitle: "Descriptive Statistics and Boxplots"
output:
  html_document: default
  pdf_document: default
---

# Learning goals

- calculate mean, median, standard deviation, and quartiles,
- obtain a five-number summary,
- create and interpret a boxplot,
- relate skewness to the choice of center.

# Example: Test scores

```{r}
scores <- c(
  31,33,35,36,40,40,41,43,45,47,
  48,51,51,52,55,56,59,70,71,74,
  76,78,84,87,87,90,93
)
```

Summary statistics:

```{r}
mean(scores)
median(scores)
sd(scores)
fivenum(scores)
summary(scores)
```

Histogram:

```{r}
hist(scores,
     main = "Test Score Distribution",
     xlab = "Score")
```

Boxplot:

```{r}
boxplot(scores,
        horizontal = TRUE,
        main = "Boxplot of Test Scores",
        xlab = "Score")
```

# Questions

1. Report the mean and median.
2. Report the five-number summary.
3. What score is the median?
4. What score is approximately the third quartile?
5. Does the distribution appear symmetric, skewed left, or skewed right?
6. Which better describes the center here: mean or median? Explain.

# Comparing two groups

```{r}
group_A <- c(72,75,77,79,81,82,83,85,86,88)
group_B <- c(55,64,68,70,74,78,82,89,94,99)

boxplot(group_A, group_B,
        names = c("Group A","Group B"),
        main = "Comparison of Two Groups")
```

# Your Turn

For each group:

1. Find the mean.
2. Find the standard deviation.
3. Find the five-number summary.
4. Compare center and spread in one or two sentences.
