---
title: "MAT 120 Computer Lab - Week 8"
subtitle: "Sampling Distributions and the Central Limit Theorem"
output:
  html_document: default
  pdf_document: default
---

# Learning goals

- understand that sample statistics vary from sample to sample,
- simulate a sampling distribution,
- connect sample size with standard error,
- use the sampling distribution of the sample mean.

# Create a population

We create a positively skewed population.

```{r}
set.seed(120)
population <- rexp(10000, rate = 1/20)

hist(population,
     main = "Population Distribution",
     xlab = "Value")
```

Population mean and standard deviation:

```{r}
mean(population)
sd(population)
```

# One random sample

```{r}
samp <- sample(population, 25)
mean(samp)
hist(samp)
```

# Sampling distribution of the mean

Generate 5000 random samples of size 25 and store each sample mean.

```{r}
sample_means_25 <- replicate(
  5000,
  mean(sample(population, 25))
)

hist(sample_means_25,
     main = "Sampling Distribution: n = 25",
     xlab = "Sample Mean")
```

Now repeat with n = 100.

```{r}
sample_means_100 <- replicate(
  5000,
  mean(sample(population, 100))
)

hist(sample_means_100,
     main = "Sampling Distribution: n = 100",
     xlab = "Sample Mean")
```

Compare standard deviations:

```{r}
sd(sample_means_25)
sd(sample_means_100)
```

# Questions

1. Is the original population symmetric or skewed?
2. What happens to the shape of the sample-mean distribution?
3. Which sampling distribution has less spread: n = 25 or n = 100?
4. Explain why using the standard-error formula.

# Probability calculation

Suppose bottle volume is normal with mean 16 oz and SD 0.6 oz.
For samples of n = 25:

```{r}
mu <- 16
sigma <- 0.6
n <- 25

se <- sigma / sqrt(n)
se

pnorm(15.8, mean = mu, sd = se)
```

# Your Turn

A population has mean 50 and standard deviation 12.

For random samples of n = 36:

1. Find the mean of the sampling distribution of x-bar.
2. Find the standard error.
3. Find P(x-bar < 47).
