---
title: "MAT 120 Computer Lab - Week 5"
subtitle: "Probability, Conditional Probability, and Independence"
output:
  html_document: default
  pdf_document: default
---

# Learning goals

- construct a two-way table,
- calculate joint and conditional probabilities,
- compare groups using conditional proportions,
- check whether two events appear independent.

# Example: Reading for fun

Suppose 500 children were surveyed.

```{r}
reading <- matrix(
  c(48,182,
    76,194),
  nrow = 2,
  byrow = FALSE,
  dimnames = list(
    Reading = c("Yes","No"),
    Gender = c("Boy","Girl")
  )
)

reading
```

Add totals.

```{r}
addmargins(reading)
```

Overall proportions:

```{r}
prop.table(reading)
```

Conditional distribution of reading status within each gender:

```{r}
prop.table(reading, margin = 2)
```

# Questions

1. What proportion of the surveyed children are boys?
2. What proportion of boys read for fun every day?
3. What proportion of girls read for fun every day?
4. Do the conditional percentages look the same?
5. Based on the sample, does reading-for-fun appear independent of gender? Explain.

# Independence check

Let:

- A = child reads for fun,
- B = child is a boy.

Compare `P(A and B)` with `P(A)P(B)`.

```{r}
p_A_and_B <- 48/500
p_A <- (48 + 76)/500
p_B <- (48 + 182)/500

p_A_and_B
p_A * p_B
```

# Your Turn

A class survey produced:

|                 | Uses tutoring | Does not use tutoring |
|-----------------|--------------:|-----------------------:|
| Passed exam     | 36            | 54                     |
| Did not pass    | 9             | 21                     |

Create the table in R, calculate conditional pass rates, and discuss whether exam result appears independent of tutoring use.
